---
title: "ChickR - ggplot2"
author: "S Bashir (ilustat.com)"
date: "26 October 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## ChickR

Welcome to ChickR

### Setup

It is assumed that the latest version of [R](www.r-project.og) is installed on your computer. It is recommended that you install and use [Rstudio IDE](https://www.rstudio.com/products/rstudio/) (integrated development environment) for the ChickR series. Other IDEs should work fine. The first step is to install some packages (*addons*/*apps*) by typing the following in the console:
```{r tidyverse, echo=TRUE, eval=FALSE}
install.packages("tidyverse")
```
It could take some time to download and install. Without going into details, we next need to *active* some of these packages as we will be using them without directly realising. So please type the following in the console:
```{r libraries, message=FALSE, results='hide'}
library(dplyr)
library(ggplot2)
```

### ChickWeight Data

The `ChickWeight` data is one of the many datasets included as part of R. The data are from a weight gain experiment for chicks. We will start by loading the data and looking at it.
```{r ChickWeightData}
data("ChickWeight")
glimpse(ChickWeight)
```
We can see that there are 4 variables and 578 observations. Each chick in the experiment is uniquely identified by the `Chick` variable (**R is case sensitive**) and they are randomly assigned to be fed one of four diets (`Diet`). Their `weight` (in grams) is measured over `Time` (from day zero to day 21). 

### ChickWeight Objective

Our objective is to investigate the effect of diet on the chieck weights over time. 

### Graphical Exploration

We will start by looking the data graphically using the `ggplot2` package. Don't worry too much about the details of the commands just try to build your own understanding.
```{r ggplot01}
ggplot(ChickWeight, aes(Time, weight)) + geom_point()
```
We see that generally the chick weights increase over time but what is the effect of diet?
```{r ggplot02}
ggplot(ChickWeight, aes(Time, weight, colour = Diet)) + geom_point()
```
It is hard to make out the effect of the diet as there are many overlapping points.
```{r ggplot03}
ggplot(ChickWeight, aes(Time, weight, colour = Diet)) + geom_jitter()
```
This looks like four hives of bees spreading out so still not easy to see what the effect of diet.
```{r ggplot04}
ggplot(ChickWeight, aes(Time, weight)) + geom_point() + facet_wrap(~Diet)
```
Perhaps it will look better with a bit of colour
```{r ggplot05}
ggplot(ChickWeight, aes(Time, weight, colour = Diet)) + geom_point() + facet_wrap(~Diet)
```
